---
title: "Leaf N and P analysis"
author: "Beni Stocker"
date: "1/6/2021"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(purrr)
library(rbeni)
library(ingestr)
library(visdat)
```

## Read data

```{r}
## original data, no covariates
df <- read_csv("~/data/LeafNP_tiandi/Global_total_leaf_N_P_Di/global_leaf_NP_total_Di_20210702_PROC_ELV.csv")

## sites (unique lon, lat, elv)
df_sites <- read_csv("~/data/LeafNP_tiandi/Global_total_leaf_N_P_Di/df_sites_leafnp_20210702.csv")

## covariates added, but incomplete
df_cov <- read_csv("~/data/LeafNP_tiandi/Global_total_leaf_N_P_Di/update_data_2021July/updated_dataset_20210429.csv")
``` 

Visualise missing.
```{r}
df_cov %>% vis_miss(warn_large_data = FALSE)
```

Fraction missing per column.
```{r}
df_cov %>% 
  summarise(across(where(is.numeric), ~(sum(is.na(.x)) / length(.x)))) %>% 
  pivot_longer(cols = 1:ncol(.), names_to = "name", values_to = "value") %>% 
  dplyr::filter(value > 0)
```

Problematic `soil_N` and `soil_P` with 88% missing. Not sure where they came from. Did you rename them, Di? 
I suspect they are from WISE and GSDE.

<!-- ```{r} -->
<!-- load("data/df_wise_20210324.RData") -->
<!-- load("data/df_gsde_20210408.RData") -->
<!-- ``` -->


## Complement dataset

### HWSD Soil

Read HWSD for each site. This reads all variables available fro HWSD. Select relevant variables in a separate step.

```{r eval=FALSE}
filn <- "data/df_hwsd_20210324.RData"
load(filn)
```

```{r}
## small number of data is missing
vis_miss(df_hwsd)
```

```{r}
sites_missing_hwsd <- df_hwsd %>% 
  dplyr::filter(is.na(T_BULK_DENSITY)) %>%  #  | is.na(T_GRAVEL) | is.na(AWC_CLASS)
  pull(sitename) %>% 
  unique()

df_sites_missing <- df_sites %>% 
  dplyr::filter(sitename %in% sites_missing_hwsd)

plot_map_simpl() +
  geom_point(data = df_sites_missing, aes(x = lon, y = lat), color = "red", size = 0.3)
```

Extract again.
```{r}
df_hwsd_compl <- ingest(
  df_sites_missing,
  source = "hwsd",
  settings = list(fil = "~/data/hwsd/HWSD_RASTER/hwsd.bil")
  )

df_hwsd_compl %>% 
  unnest(data) %>% 
  vis_miss()
```

The problem is that the data is just missing for these locations, there is no way to resolve it on "ingestr-side". The fraction of missing data is 0.6%. That's acceptable.

### WISE30sec Soil

Define all variables to be extracted from WISE30sec

```{r}
load("data/df_wise_20210324.RData")
```

Practically no missing data (<0.1%).
```{r}
vis_miss(df_wise)
```

### GSDE

```{r}
load("data/df_gsde_20210408.RData")
```

Lots of missing data!
```{r}
vis_miss(df_gsde)
```

Map missing data. No particular pattern.
```{r}
sites_missing_gsde <- df_gsde %>% 
  dplyr::filter(is.na(PBR)) %>%  #  | is.na(T_GRAVEL) | is.na(AWC_CLASS)
  pull(sitename) %>% 
  unique()

df_sites_missing <- df_sites %>% 
  dplyr::filter(sitename %in% sites_missing_gsde)

plot_map_simpl() +
  geom_point(data = df_sites_missing, aes(x = lon, y = lat), color = "red", size = 0.3)
```

Try extracting again.
```{r}
settings_gsde <- list(varnam =  c("PBR"), layer = 1:4)    # , "PHO", "TP", "TK", "TS"

df_gsde_compl <- ingest(
  df_sites_missing %>% slice(1:3),
  source    = "gsde",
  settings  = settings_gsde,
  dir       = "~/data/soil/shangguan"
  ) %>% 
  unnest(data)
```


### Topography index

0.26% missing. Acceptable.

```{r}
load("data/df_gti_20210324.RData")
vis_miss(df_gti)
```

### CO2

No missing.

```{r}
load("data/df_co2_20210324.RData")
vis_miss(df_co2)
```

### Nitrogen deposition

No missing. 

```{r}
load("data/df_ndep_20210324.RData")
vis_miss(df_ndep_agg)
```

### P-model outputs and climate indeces

The script `rscript_ingest_run_rsofun.R` is used to collect forcing data and run P-model simulations for all sites (N = 2149). Outputs are aggregated by site (annual mean of Vcmax25 and Jmax25, weighted by daily GPP). This also runs the function `R/calc_climate_index.R` on collected climate forcing data to return climate indices (some of which are complementary to what is collected directly from WorldClim). The output is on Euler and is read here.

0.2% missing. Acceptable.

```{r}
read_pmodel <- function(path){
  load(path)
  return(df_pmodel)
}
df_pmodel <- purrr::map_dfr(as.list(list.files("data", pattern = "df_pmodel.*.RData", full.names = TRUE)),
                            ~read_pmodel(.))

save(df_pmodel, file = "data/df_pmodel.RData")

vis_miss(df_pmodel)
```

Large number of data missing for alpha. Unclear why. Resolve it.


## Combine all data

```{r}
load("data/df_ndep_20210324.RData") # loads df_ndep_agg
load("data/df_co2_20210324.RData") # df_co2
load("data/df_gti_20210324.RData") # loads df_gti
load("data/df_gsde_20210408.RData") # df_gsde
load("data/df_wise_20210324.RData") # df_wise
load("data/df_hwsd_20210324.RData") # df_hwsd

df <- read_csv("~/data/LeafNP_tiandi/Global_total_leaf_N_P_Di/global_leaf_NP_total_Di_20210702_PROC_ELV.csv")

df_my <- df %>% 
  dplyr::select(-Record_ID, -soil_N, -soil_P) %>% 
  left_join(df_wise, by = "sitename") %>% 
  left_join(df_hwsd, by = "sitename") %>% 
  left_join(df_gti %>% dplyr::select(-year_start, -year_end, -elv, -lon, -lat), by = "sitename") %>% 
  left_join(df_ndep_agg, by = "sitename") %>% 
  left_join(df_co2, by = "sitename") %>% 
  left_join(df_pmodel, by = "sitename")

## data prepared by di
df_di <- read_csv("~/data/LeafNP_tiandi/Global_total_leaf_N_P_Di/soil_property_extraction_20210323/global_leaf_NP_with_soil_property_from_HWSD_WISE_GSDE_23032021.csv")

## add my extractions to di's soil stuff
df_di <- df_di %>% 
  left_join(df_gti %>% dplyr::select(-year_start, -year_end, -elv, -lon, -lat), by = "sitename") %>% 
  left_join(df_ndep_agg, by = "sitename") %>% 
  left_join(df_co2, by = "sitename") %>% 
  left_join(df_pmodel, by = "sitename")

write_csv(df_di, "~/data/LeafNP_tiandi/Global_total_leaf_N_P_Di/soil_property_extraction_20210323/global_leaf_NP_with_soil_property_from_HWSD_WISE_GSDE_Pmodel_Ndep_GTI_CO2_25032021.csv")  

dim(df_my)
dim(df_di)
dim(df)

names(df_my)

## are my extractions and the one of di the same?
## yes
df_my %>% 
  dplyr::select(sitename, T_CACO3_beni = T_CACO3) %>% 
  left_join(dplyr::select(df_di, sitename, T_CACO3_di = T_CACO3), by = "sitename") %>% 
  rbeni::analyse_modobs2("T_CACO3_beni", "T_CACO3_di", type = "hex")

## yes
df_my %>% 
  dplyr::select(sitename, T_PH_H2O_beni = T_PH_H2O) %>% 
  left_join(dplyr::select(df_di, sitename, T_PH_H2O_di = T_PH_H2O), by = "sitename") %>% 
  rbeni::analyse_modobs2("T_PH_H2O_beni", "T_PH_H2O_di", type = "hex")

df_my %>% 
  dplyr::select(sitename, CNrt_beni = CNrt) %>% 
  left_join(dplyr::select(df_di, sitename, CNrt_di = CNrt), by = "sitename") %>% 
  rbeni::analyse_modobs2("CNrt_beni", "CNrt_di", type = "hex")
```

### Extracting from global NetCDF files

This is a code example and is to be used for extracting AI, alpha, and CWDX40 (see [here](https://www.notion.so/computationales/4e2e0ca6b5d3497894c64bed71a4937d?v=0ee5d81cd0fe42ac8ac1ed0a938b3380)).

Use the following files:

-   `~/data/mct_data/cwdx40.nc` for CWDX40.
-   `~/data/sofun_outputs/global_FULL_MODIS-C006_MOD15A2_v3.4/global_FULL_MODIS-C006_MOD15A2_v3.4.a.alpha_MEANANN.nc` for alpha.

```{r}
library(raster)

filn <- "~/data/mct_data/cwdx40.nc"
siteinfo <- tibble(sitename = "id1", lon = 100, lat = 50)  # example for required dataframe structure

## using raster package
rasta <- raster::raster(filn)  # use raster::brick() if data has three dimensions
df <- raster::extract(
    rasta,
    sp::SpatialPoints(dplyr::select(siteinfo, lon, lat)), # , proj4string = sta@crs
    sp = TRUE
    ) %>%
  as_tibble()

## or simply using rbeni
df <- rbeni::extract_pointdata_allsites(filn, df_lonlat = siteinfo)
```
