---
title: "Leaf N and P analysis"
author: "Beni Stocker"
date: "1/6/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rbeni)
library(ingestr)
```

## Read data

```{r}
df <- read_csv("~/data/LeafNP_tiandi/SLA_updated_leafNP_Di/Global_N_P_dataset_20210110_final.csv") %>%  #, col_types = "idddcccdddddccccccc") %>% 
  mutate(lon = round(Lon_Di_check_final, digits = 4),
         lat = round(Lat_Di_check_final, digits = 4),
         elv = round(Alt_Di_check_final, digits = -1)) %>% 
  rename(year_start = Sampling_Year_start,
         year_end = Sampling_Year_end) %>% 
  
  ## to be done: make sampling year info machine readable: either in the form of "9999" or "8888_9999"
  separate(Sampling_Month, sep = "_", c("month_start", "month_end")) 
  # rowwise() %>% 
  # mutate(year_mid = mean(year_start, year_end, na.rm = TRUE))
```

### Complement missing elevation

```{r}
df_tmp <- df %>% 
  dplyr::select(lon, lat, elv) %>% 
  dplyr::filter(is.na(elv)) %>% 
  dplyr::select(lon, lat) %>% 
  distinct() %>% 
  mutate(sitename = paste0("i", 1:n()))

df_etopo <- ingest(
  df_tmp,
  source = "etopo1",
  dir = "~/data/etopo/"  # adjust this with your local path
  ) %>% 
  unnest(data) %>% 
  right_join(df_tmp, by = "sitename") %>% 
  rename(elv_etopo = elv)
  
df <- df %>% 
  left_join(df_etopo, by = c("lon", "lat")) %>% 
  rowwise() %>% 
  mutate(elv = ifelse(is.na(elv), elv_etopo, elv)) %>% 
  dplyr::select(-sitename, -elv_etopo)
```

## Identify sites

Identify sites (unique lon, lat, elevation)

```{r}
df_sites <- df %>%
  dplyr::select(lon, lat, elv) %>% 
  distinct() %>% 
  mutate(sitename = paste0("i_", as.character(lon), "_", as.character(lat), "_", as.character(elv)))

df <- df %>% 
  left_join(df_sites, by = c("lon", "lat", "elv"))

## determine start and end year for each site based on available measurements
df_sites <- df %>% dplyr::select(sitename, year_start, year_end) %>% 
  group_by(sitename) %>% 
  summarise(year_start = min(year_start, na.rm = TRUE),
            year_end   = max(year_end, na.rm = TRUE)) %>% 
  right_join(df_sites, by = "sitename")

write_csv(df_sites, file = "data/df_sites.csv")
```

Sites on a global map

```{r}
plot_map_simpl() +
  geom_point(data = df_sites, aes(x = lon, y = lat), color = "red", size = 0.3)
```

## Identify cells

Based on half-degree global grid. ACTUALLY NOT NEEDED.

```{r eval=FALSE}
## bin to half-degree gridcells for determining climate forcing
dlon <- 0.5
dlat <- 0.5
lon_breaks <- seq(from = floor(min(df_sites$lon)), to = ceiling(max(df_sites$lon)), by = dlon)
lat_breaks <- seq(from = floor(min(df_sites$lat)), to = ceiling(max(df_sites$lat)), by = dlat)

df_sites <- df_sites %>%
  ungroup() %>% 
  mutate(ilon = cut(lon, 
                    breaks = lon_breaks
                    ),
         ilat = cut(lat, 
                    breaks = lat_breaks
                    )
         ) %>% 
  mutate(lon_lower = as.numeric( sub("\\((.+),.*", "\\1", ilon)),
         lon_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilon) ),
         lat_lower = as.numeric( sub("\\((.+),.*", "\\1", ilat) ),
         lat_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilat) )
         ) %>% 
  mutate(lon_mid = (lon_lower + lon_upper)/2,
         lat_mid = (lat_lower + lat_upper)/2) %>% 
  
  ## create cell name to associate with climate input
  mutate(cellname = paste0("icell_", as.character(lon_mid), "_", as.character(lat_mid))) %>% 
  dplyr::select(-ilon, -ilat, -lon_lower, -lon_upper, -lat_lower, -lat_upper)

df_cells <- df_sites %>% 
  dplyr::select(cellname, lon_mid, lat_mid) %>% 
  distinct()
```

## Complement dataset

*Look into:*

-   Complement dataset with data points where only Nmass or only Pmass is available (no strict requirement to have paired Nmass and Pmass for this analysis).
-   Information about species identity, family, genus

### Climate indeces

This is how it can be done.

```{r}
## ddf_watch must be a flat data frame (not nested)
## functions to calculate all climate indexes are implemented here:
source("R/calc_climate_index.R")

## test: use just data for two sites
testcells <- unique(ddf_watch$sitename)[1:2]

df_climate_ind <- ddf_watch %>% 
  dplyr::filter(sitename %in% testcells) %>% 
  group_by(sitename) %>% 
  nest() %>% 
  mutate(mat = purrr::map(data, ~calc_climate_index_mat(.)),
         matgs = purrr::map(data, ~calc_climate_index_matgs(., temp_base = 5.0)),
         tmonthmin = purrr::map(data, ~calc_climate_index_tmonthmin(.)),
         tmonthmax = purrr::map(data, ~calc_climate_index_tmonthmax(.)),
         ndaysgs = purrr::map(data, ~calc_climate_index_ndaysgs(., temp_base = 5.0)),
         mai = purrr::map(data, ~calc_climate_index_mai(.)),
         maigs = purrr::map(data, ~calc_climate_index_maigs(., temp_base = 5.0)),
         map = purrr::map(data, ~calc_climate_index_map(.)),
         pmonthmin = purrr::map(data, ~calc_climate_index_pmonthmin(.)),
         mapgs = purrr::map(data, ~calc_climate_index_mapgs(., temp_base = 5.0)),
         mavgs = purrr::map(data, ~calc_climate_index_mavgs(., temp_base = 5.0)),
         mav = purrr::map(data, ~calc_climate_index_mav(.))) %>% 
  dplyr::select(-data) %>% 
  unnest(c(mat, matgs, tmonthmin, tmonthmax, ndaysgs, mai, maigs, map, pmonthmin, mapgs, mavgs, mav))
```

### HWSD Soil

Read HWSD for each site. This reads all variables available fro HWSD. Select relevant variables in a separate step.

```{r eval=FALSE}
filn <- "data/df_hwsd.RData"
if (!file.exists(filn)){
  df_hwsd <- ingest(
    df_sites,
    source = "hwsd",
    settings = list(fil = "~/data/hwsd/HWSD_RASTER/hwsd.bil")
    )
  
  ## for some it got info for multiple soil types - take only first ones => TEST INFLUENCE OF THIS
  df_hwsd <- df_hwsd %>% 
    mutate(n_row = purrr::map_int(data, ~nrow(.))) %>% 
    mutate(data = purrr::map(data, ~slice(., 1)))
  
  save(df_hwsd, file = filn)
  
} else {
  load(filn)
}
```

Combine datasets.

```{r}
df <- df %>% 
  left_join(
    df_hwsd %>% 
      unnest(data)
  )
```

### SoilGrids

Define all variables to be extracted from SoilGrids

```{r}
settings_soilgrids <- get_settings_soilgrids(
  varnam = c("nitrogen", "cec"), # xxx please adjust
  layer = 1:3                    # xxx please adjust
  )
```

Get and save data.

```{r}
df_soilgrids <- ingest(
  df_sites,
  source    = "soilgrids",
  settings  = settings_soilgrids
  ) %>% 
  unnsest(data)  # to get a normal flat table

save(df_soilgrids, file = "data/df_soilgrids.RData")
```

### WISE30sec Soil

Define all variables to be extracted from WISE30sec

```{r}
settings_wise <- get_settings_wise(
  varnam = c("CNrt", "CECS"),  # xxx please adjust
  layer = 1:3                  # xxx please adjust
  )
```

Get and save data.

```{r}
df_wise <- ingest(
  df_sites,
  source    = "wise",
  settings  = settings_wise,
  dir       = "~/data/soil/wise"
  ) %>% 
  unnsest(data)  # to get a normal flat table

dave(df_wise, file = "data/df_wise.RData")
```

### GSDE Soil XXX

Define all variables to be extracted from GSDE. This is not using ingestr.

```{r}
settings_wise <- get_settings_gsde(
  varnam = c("CNrt", "CECS"),  # xxx please adjust
  layer = 1:3                  # xxx please adjust
  )
```

Get data.

```{r}
df_wise <- ingest(
  df_sites,
  source    = "wise",
  settings  = settings_wise,
  dir       = "~/data/soil/wise"
  )
```

### Topography index

```{r}
filn <- "data/df_gti.RData"
if (!file.exists(filn)){
  rasta <- raster::raster("~/data/gti_marthews/ga2.nc")
  df_gti <- raster::extract(rasta, sp::SpatialPoints(dplyr::select(df_sites, lon, lat) %>% distinct()), sp = TRUE) %>%
    as_tibble() %>% 
    rename(gti = GDAL.Band.Number.1)
  save(df_gti, file = filn)
} else {
  load(filn)
}

## add to data 
df <- df_gti%>% 
      right_join(
        df,
        by = c("lon", "lat")
      )
```

### CO2

Read CO2 level as the global mean and a mean over all the years over which sampling was done. This means that if we have replicate measurements for a given species at a given site, the measurements are given the same CO2 level. This assumes that replicates are not intended to cover a time period over which the environment changes.

```{r}
df_co2 <- ingest(
  df_sites,
  source  = "co2_mlo",
  verbose = FALSE
  ) %>% 
  
  ## aggregate over years per site
  mutate(co2 = purrr::map_dbl(data, ~{summarise(., co2 = mean(co2)) %>% pull(co2)})) %>% 
  dplyr::select(-data)

## add to data
df <- df_co2 %>% 
  right_join(df, by = "sitename")
```

### Nitrogen deposition

Read N deposition from Lamarque et al. (2011) as NOx and NOy in gN m$^{-2}$ yr$^{-1}$ for respective year and location (based on half-degree input file).

```{r}
filn <- "data/df_ndep.RData"
if (!file.exists(filn)){
  df_ndep <- ingest(
    df_sites %>% 
      ## for some we don't have data beyond 2009
      rowwise() %>% 
      mutate(year_start = min(year_start, 2009),
             year_end = min(year_end, 2009)),
    source    = "ndep",
    timescale = "y",
    dir       = "~/data/ndep_lamarque/",
    verbose   = FALSE
    )
  save(df_ndep, file = filn)
} else {
  load(filn)
}

## take sum of noy and nhx and aggregate over years
df_ndep_agg <- df_ndep %>% 
  mutate(data = purrr::map(data, ~mutate(., ndep = noy + nhx))) %>% 
  mutate(data = purrr::map(data, ~summarise(., ndep = mean(ndep, na.rm = TRUE)))) %>% 
  unnest(data)

## add to data
df <- df %>% 
  left_join(df_ndep_agg, by = "sitename")
```
