---
title: "P-model simulations for leafNP"
author: "Beni"
date: "1/15/2021"
output: html_document
---

```{r}
#knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ingestr)
library(rsofun)
```

The challenge is to use very site specific forcing data for each site and the data load required for generating full forcing and output time series for ~30,000 sites. In the end, we only need one value per site (Vcmax25). Let's therefore wrap collecting forcing data, running the model, and aggregating outputs to the required single value into a one function. A further challenge is that if all original forcing files have to be opened and read for each site, it takes forever. We cannot do this for all sites at once becaause of RAM memory overload. An optimal middle ground may be to do split this up into chunks, each comprising 100 sites. That makes ~300 chunks. That is 300 simulations in parallel with manageable memory requirements. Let's do that.

Define a function wrapping the ingest and the pmodel steps for a site of sites, given their lon and lat (and elv and years range), provided as a data frame. The following is an example for such a site set. This is implemented by `ingest_run_rsofun()`.
```{r}
siteinfo <- tibble(
  sitename = c("site_rsofun_demo_1", "site_rsofun_demo_2", "site_rsofun_demo_3"),
  lon = c(100.05, 100.45, 90),
  lat = c(50, 50, 50),
  elv = c(1000, 250, 2500),
  year_start = c(1960, 2001, 2005),
  year_end = c(2005, 2006, 2007),
  whc = 200
)

df <- ingest_run_rsofun(siteinfo)
```



---

# OLD

## Site meta information

Read the site meta information data frame, created in `leafnp.Rmd`. For site ensemble P-model simulations, it must contain the following columns:

- `sitename`
- `lon`
- `lat`
- `elv`
- `year_start`
- `year_end`

```{r}
df_sites <- read_csv("data/df_sites.csv")
```

# Simulation by sites

In this case, we're extracting climate forcing data for each site (irrespective of whether multiple sites are located within a single gridcell of the global forcing data sets).


# Simulation by gridcells

Since we're extracting forcing data from global files provided at half-degree resolution, we want to avoid extracting data for two sites separately if they are located within the same gridcell. This has the advantage of reducing the number of sites, but comes at the disadvantage of resolving within-gridcell differences between different sites' climate forcing. This may be negligible, except for the conversion of specific humidity to vapour pressure deficit, which takes elevation as input. Note that the explicit elevation-dependence (i.e. atmospheric pressure-dependence) of photosynthetic processes comes in through using the site's elevation as an argument to the model call.

## Determine gridcells

Therefore, determine the unique gridcells.

```{r}
## bin
dlon <- 0.5
dlat <- 0.5
lon_breaks <- seq(from = floor(min(df_sites$lon)), to = ceiling(max(df_sites$lon)), by = dlon)
lat_breaks <- seq(from = floor(min(df_sites$lat)), to = ceiling(max(df_sites$lat)), by = dlat)

df_sites <- df_sites %>%
  ungroup() %>% 
  mutate(ilon = cut(lon, 
                    breaks = lon_breaks
                    ),
         ilat = cut(lat, 
                    breaks = lat_breaks
                    )
         ) %>% 
  mutate(lon_lower = as.numeric( sub("\\((.+),.*", "\\1", ilon)),
         lon_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilon) ),
         lat_lower = as.numeric( sub("\\((.+),.*", "\\1", ilat) ),
         lat_upper = as.numeric( sub("[^,]*,([^]]*)\\]", "\\1", ilat) )
         ) %>% 
  mutate(lon_mid = (lon_lower + lon_upper)/2,
         lat_mid = (lat_lower + lat_upper)/2) %>% 
  dplyr::select(-ilon, -ilat, -lon_lower, -lon_upper, -lat_lower, -lat_upper) %>% 
  mutate(cellname = paste0("icell_", lon_mid, "_", lat_mid))

## save again, now with lon_mid, lat_mid
save(df_sites, file = "data/df_sites.RData")

df_cells <- df_sites %>% 
  dplyr::select(sitename = cellname, lon = lon_mid, lat = lat_mid) %>% 
  distinct() %>% 

  ## set for reading all available years, subset to required years for each site later
  mutate(year_start = 1979, year_end = 2018)

write_csv(df_cells, path = "data/df_cells.csv")
```

## Read forcing data

### Climate

Use the ingestr package to read WATCH-WFDEI climate data as daily time series for each cell that contains at least one site. 

Actually it's best to do this on the Euler cluster. Submit it as a job with the submission file `submit_get_watch_cru.sh` (a Bash script), running an R script (`rscript_get_watch_cru.R`) that contains the same code as the chunk below.

```{r eval=FALSE}
df_cells <- read_csv("data/df_cells.csv")

ddf_watch <- ingest(
  siteinfo = df_cells,
  source    = "watch_wfdei",
  getvars   = c("temp", "prec", "ppfd", "vpd"),
  dir       = "~/data/watch_wfdei/"  # adjust this with your local path
)
save(ddf_watch, file = "data/ddf_watch.RData")

## make flat and write to file
ddf_watch %>% 
  unnest(data) %>% 
  write_csv(file =  "data/ddf_watch.csv")

ddf_cru <- ingest(
  siteinfo = df_cells,
  source    = "cru",
  getvars   = "ccov",
  dir       = "~/data/cru/ts_4.01/"  # adjust this with your local path
)
save(ddf_cru, file = "data/ddf_cru.RData")

## make flat and write to file
ddf_cru %>% 
  unnest(data) %>% 
  write_csv(file =  "data/ddf_cru.csv")
```

Combine the two data frames for P-model forcing.

```{r eval=FALSE}
load("data/ddf_watch.RData")
load("data/ddf_cru.RData")

ddf_meteo <- ddf_watch %>% 
  tidyr::unnest(data) %>% 
  left_join(
    ddf_cru %>% 
      tidyr::unnest(data),
    by = c("sitename", "date")
  ) %>% 
  group_by(sitename) %>% 
  tidyr::nest()
```

### CO2

```{r eval=FALSE}
df_co2 <- read_csv("~/data/co2/cCO2_rcp85_const850-1765.csv")
```


## Set P-model parameters

```{r}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = TRUE,
  tempstress         = TRUE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
	)

## calibrated parameters for v3.0 (see https://rpubs.com/stineb/rsofun_benchmark_v30)
params_modl <- list(
	kphio           = 0.09423773,
	soilm_par_a     = 0.33349283,
	soilm_par_b     = 1.45602286,
	vpdstress_par_a = 9999,
	vpdstress_par_b = 9999,
	vpdstress_par_m = 9999
	)

df_soiltexture <- bind_rows(
  top    = tibble(layer = "top",    fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1),
  bottom = tibble(layer = "bottom", fsand = 0.4, fclay = 0.3, forg = 0.1, fgravel = 0.1)
)
```

## Prepare drivers for P-model

Run `prepare_setup_sofun()` to define the simulation settings that contain all the information specified by the two steps above (meta info, and simulation parameters), global simulation parameters are wrapped inside an additional column `params_siml`, added to the site meta info dataframe.
```{r}
df_drivers <- prepare_setup_sofun(siteinfo = df_sites, params_siml = params_siml)
```

Nest site info
```{r}
df_drivers <- df_drivers %>% 
  dplyr::select(sitename, lon, lat, elv) %>% 
  mutate(c4 = FALSE, whc = 240) %>% 
  group_by(sitename) %>% 
  nest() %>% 
  rename(siteinfo = data) %>% 
  left_join(df_drivers %>% 
              dplyr::select(sitename, params_siml, cellname),
            by = "sitename")
```

Add climate forcing and subset time series based on start and end year for this site.
```{r}
df_co2 <- read_csv("~/data/co2/cCO2_rcp85_const850-1765.csv")

subset_forcing <- function(df_forcing, params_siml){
  
  ## get mean seasonality in forcing to be used for all years before 1979
  df_forcing_meandoy <- df_forcing %>% 
    # dplyr::filter(year(date) %in% 1979:1988) %>% 
    ungroup() %>% 
    mutate(doy = lubridate::yday(date)) %>% 
    group_by(doy) %>% 
    summarise_if(is.numeric, mean, na.rm = TRUE)
  
  ## initialise time series for required years
  ddf <- init_dates_dataframe(
    yrstart = params_siml$firstyeartrend[1],
    yrend = params_siml$firstyeartrend[1] + params_siml$nyeartrend[1] - 1,
    noleap = TRUE
    )

  ## add available forcing data to given time series 
  ddf <- ddf %>% 
    left_join(df_forcing, by = "date") %>% 
    
    ## add mean seasonal cycle
    mutate(doy = lubridate::yday(date)) %>% 
    left_join(df_forcing_meandoy %>% 
                rename(temp_doy = temp, patm_doy = patm, qair_doy = qair, vpd_doy = vpd,
                       ppfd_doy = ppfd, prec_doy = prec, ccov_doy = ccov),
                by = "doy") %>% 
    
    ## fill missing with mean seasonal cycle
    rowwise() %>% 
    mutate(temp = ifelse(is.na(temp), temp_doy, temp),
           vpd  = ifelse(is.na(vpd),  vpd_doy,  vpd),
           patm = ifelse(is.na(patm), patm_doy, patm),
           ppfd = ifelse(is.na(ppfd), ppfd_doy, ppfd),
           prec = ifelse(is.na(prec), prec_doy, prec),
           ccov = ifelse(is.na(ccov), ccov_doy, ccov)
    ) %>% 
    
    ## merge co2 data in it
    mutate(year = lubridate::year(date)) %>% 
    left_join(df_co2, by = "year") %>% 
    
    ## remove days in leap years
    dplyr::filter(!(month(date)==2 & mday(date) == 29))
  
  return(ddf)
}

df_drivers <- df_drivers %>% 
  left_join(ddf_meteo %>% 
              rename(cellname = sitename), 
            by = "cellname") %>% 
  mutate(forcing = purrr::map2(forcing, params_siml, ~subset_forcing(.x, .y)))
```

Add soil texture data frame (for completeness, but not actually used).
```{r}
df_drivers <- df_soiltexture %>%
  dplyr::mutate(tmp = 1) %>% 
  group_by(tmp) %>% 
  nest() %>% 
  rename(df_soiltexture = data) %>% 
  right_join(
    df_drivers %>% 
      mutate(tmp = 1),
    by = "tmp"
  ) %>% 
  dplyr::select(-tmp)
```

Right order of columns for `pmap`.
```{r}
df_drivers <- df_drivers %>% 
  dplyr::select(sitename, params_siml, siteinfo, df_soiltexture, forcing)
```

## Run P-model

This is a tiny bit more complicated than in the simplest case (described[here](https://stineb.github.io/rsofun/articles/example_pmodel.html#run-the-model)), because we're combining site meta information from `df_sites` with climate forcing by gridcell, based on the gridcell ID.
```{r}
df_pmodel <- runread_pmodel_f(
  df_drivers,
  params_modl = params_modl,
  makecheck = TRUE,
  parallel = FALSE
  )

save(df_pmodel, file = "data/df_pmodel.RData")

df_pmodel_flat <- df_pmodel %>% 
  unnest(out)
write_csv(df_pmodel_flat, path = "data/df_pmodel_flat.csv")
```
