---
title: "Analyse order of predictors in fitting models"
author: "Beni Stocker"
date: "2022-07-28"
output: pdf_document
---

This is based on the script shared by Bernhard Schmid and contained also this repo (`analysis/TianDiMay23_BS.R`).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)

library(bestNormalize)
library(nlme)
# library(lme4)
library(lmerTest)
library(effects)
library(ggeffects)
library(performance)

df_model_fitting <- tibble()
overwrite <- FALSE
```

## Read and pre-process data

Read full species-level data
```{r}
df <- read_csv("~/data/LeafNP_tiandi/Global_total_leaf_N_P_Di/leafnp_data_covariates_20210702.csv") %>% 
  mutate(grass = tree_shrub_Herb == "H")  
```

Define predictor sets.
```{r}
trgts <- c("leafN", "leafP", "LeafNP")

## predictors excluding PHO, and TS (too many missing)
preds <- c("elv", "mat", "matgs", "tmonthmin", "tmonthmax", "ndaysgs", "mai", "maigs", "map", "pmonthmin", "mapgs", "mavgs", "mav", "alpha", "vcmax25", "jmax25", "gs_accl", "aet", "ai", "cwdx80", "gti", "ndep", "co2", "T_BULK_DENSITY", "AWC_CLASS", "T_CLAY", "T_SILT", "T_SAND", "T_GRAVEL", "T_PH_H2O", "T_TEB", "T_BS", "T_CEC_SOIL", "T_CEC_CLAY", "T_ECE", "T_ESP", "T_CACO3", "T_OC", "ORGC", "TOTN", "CNrt", "ALSA", "PBR", "TP", "TK")
```

Filter to use only data from species that were recorded in at least five different sites.
```{r}
# yields 398 species
use_species <- df %>% 
  dplyr::select(sitename, Species) %>% 
  distinct() %>% 
  group_by(Species) %>% 
  summarise(n = n()) %>% 
  dplyr::filter(n >= 5) %>% 
  pull(Species)
```

Additionally, analogous to how it's done for the trait gradient analysis, use data only for sites where at least M species are present.
```{r}
use_sites <- df %>% 
  dplyr::select(sitename, Species) %>% 
  distinct() %>% 
  group_by(sitename) %>% 
  summarise(n = n()) %>% 
  dplyr::filter(n >= 5) %>% 
  pull(sitename)
```

Apply filtering.
```{r}
df <- df %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites)
```

Prepare date for linear mixed effects modelling. Do data pre-processing as Di Tian implemented it. Transformations are required, but not for Random Forest modelling. Therefore, define separate data objects for the two model types.

Make factors and numeric. No capping to positive numbers done here (as opposed to Di Tian's).
```{r}
df_lmm <- df %>%
  # drop_na() %>%
  rename( 
    SITE = sitename,
    SP = Species, 
    FG = FunGroups, 
    TG = tree_shrub_Herb, 
    FGID = Dc_Db_Ec_Eb_Hf_Hg,
    FA = Family, 
    GE = Genus, 
    ID = id
    ) %>%
  mutate(across(all_of(c("SITE", "FG", "FGID", "TG", "FA", "GE", "SP", "ID")), factor)) %>%
  mutate(across(all_of(preds), as.numeric))
```

<!-- Apply Yeo-Johnson transformation to make predictors more normally distributed. -->
<!-- ```{r} -->
<!-- predict_yeojohnson <- function(x){ -->
<!--   # x[x < 0] <- 0   # How Di did it -->
<!--   predict( yeojohnson( x ) ) -->
<!-- } -->

<!-- df_lmm <- df_lmm %>% -->
<!--   mutate(across(all_of(preds), predict_yeojohnson)) -->
<!-- ``` -->

## ANOVA for differet model specifications

### LM with species first

```{r}
# individual environmental factors
lmm2 <- lm(
  terms(log(leafN) ~ SP + (ALSA + mav + elv + co2 + tmonthmin + mai + ndep),
        keep.order = TRUE),
  data = df_lmm |>
    drop_na())

df_ss2 <- as_tibble(anova(lmm2), rownames = "Predictor") |> 
  select(Predictor, `Species first`=`Sum Sq`)

df_ss2_sum <- df_ss2 |> 
  summarise(`Species first` = sum(`Species first`))

df_ss2 <- df_ss2 |> 
  mutate(`Species first (%)` = `Species first`/df_ss2_sum$`Species first`)
```

### LM with species last

```{r}
# individual environmental factors
lmm3 <- lm(
  terms(log(leafN) ~ (ALSA + mav + elv + co2 + tmonthmin + mai + ndep) + SP,
        keep.order = TRUE),
  data = df_lmm |>
    drop_na())

df_ss3 <- as_tibble(anova(lmm3), rownames = "Predictor") |> 
  select(Predictor, `Species last`=`Sum Sq`)

df_ss3_sum <- df_ss3 |> 
  summarise(`Species last` = sum(`Species last`))

df_ss3 <- df_ss3 |> 
  mutate(`Species last (%)` = `Species last`/df_ss3_sum$`Species last`)
```

### LMM with species as random

```{r}
# lmer package
# lmm4 <- lme(
# 
#   log(leafN) ~ ALSA + mav + elv + co2 + tmonthmin + mai + ndep,
#   random = ~1|SP,
#   data = df_lmm |> 
#     drop_na()
# 
#   )

# lme4 package, individual environmental factors
lmm4 <- lme4::lmer(
  log(leafN) ~ ALSA + mav + elv + co2 + tmonthmin + mai + ndep +
    (1|SP),
  data = df_lmm |>
    drop_na()
  )

df_ss4 <- as_tibble(anova(lmm4), rownames = "Predictor") |>
  select(Predictor, `Species as random`=`Sum Sq`)

df_ss4_sum <- df_ss4 |>
  summarise(`Species as random` = sum(`Species as random`))

df_ss4 <- df_ss4 |>
  mutate(`Species as random (%)` = `Species as random`/df_ss4_sum$`Species as random`)

# # lme4 package, all environmental factors wrapped into site
# lmm4 <- lme4::lmer(
#   log(leafN) ~ 1 + (1|SP) + (1|SITE),
#   data = df_lmm |>
#     drop_na()
#   )
# 
# summary(lmm4)
```

Sum of squares for each predictor in the three different models.
```{r}
df_ss2 |> 
  left_join(
    df_ss3,
    by = "Predictor"
  ) |> 
  left_join(
    df_ss4,
    by = "Predictor"
  ) |> 
  mutate(
    `Species first` = format(`Species first`, digits = 2),
    `Species last` = format(`Species last`, digits = 2),
    `Species as random` = format(`Species as random`, digits = 2)
  ) |> 
  select(c(1,2,4,6)) |> 
  knitr::kable()
```

## Venn diagram for leaf N

Here, site summarises all possible environmental factors.

### Species first

```{r}
mod1 <- lm(
  terms(log(leafN) ~ SP + SITE,
        keep.order = TRUE),
  data = df_lmm |>
    drop_na())

df1 <- as_tibble(anova(mod1), rownames = "Predictor") |>
  select(Predictor, species_first=`Sum Sq`)

df1_sum <- df1 |>
  summarise(species_first = sum(species_first))

df1 <- df1 |>
  mutate(species_first_perc = species_first/df1_sum$species_first)
```

### Species last

```{r}
mod2 <- lm(
  terms(log(leafN) ~ SITE + SP,
        keep.order = TRUE),
  data = df_lmm |>
    drop_na())

df2 <- as_tibble(anova(mod2), rownames = "Predictor") |>
  select(Predictor, species_last=`Sum Sq`)

df2_sum <- df2 |>
  summarise(species_last = sum(species_last))

df2 <- df2 |>
  mutate(species_last_perc = species_last/df2_sum$species_last)
```

```{r}
df_site_species <- df1 |> 
  left_join(
    df2,
    by = "Predictor"
  )

df_site_species |> knitr::kable()
```

### Plot

```{r}
df_shared <- tibble(
  species = df_site_species |> 
    select(Predictor, species_last) |> 
    filter(Predictor == "SP") |> 
    pull(species_last),
  site = df_site_species |> 
    select(Predictor, species_first) |> 
    filter(Predictor == "SITE") |> 
    pull(species_first),
  species_plus = df_site_species |> 
    select(Predictor, species_first) |> 
    filter(Predictor == "SP") |> 
    pull(species_first),
  site_plus = df_site_species |> 
    select(Predictor, species_last) |> 
    filter(Predictor == "SITE") |> 
    pull(species_last)
) |> 
  mutate(
    shared_a = species_plus - species,
    shared_b = site_plus - site
  )

total <- df_shared$species_plus + df_shared$site

df_shared_perc <- df_shared |> 
  mutate(across(1:6, ~ 100*(. / total)))

df_venn <- tibble(x = c(0,7),
                  y = c(0,0),
                  r = c(sqrt(df_shared$site_plus), sqrt(df_shared$species_plus)),
                  id = c("site", "species")
                  )

gg_n <- ggplot(df_venn, 
             aes(x0 = x, y0 = y, r = r, fill = id)) +
  ggforce::geom_circle(alpha = .5, show.legend = FALSE ) +
  coord_fixed() +
  scale_fill_manual(values=c("#29a274ff", "#777055ff")) +
  theme_void() +
  annotate("text", x = -8 , y = 0, label = paste0("Site \n ", 
                          format(df_shared_perc$site, digits = 2), 
                          "%")) +
  annotate("text", x = 3 , y = 0, 
           label = paste0("Shared \n ", 
                          format(df_shared_perc$shared_a, digits = 2), 
                          "%")) +
  annotate("text", x = 17 , y = 0, 
           label = paste0("Species \n ", 
                          format(df_shared_perc$species, digits = 2), 
                          "%")) +
  labs(title = "Leaf N")
```

## Venn diagram for leaf P

Here, site summarises all possible environmental factors.

### Species first

```{r}
mod1 <- lm(
  terms(log(leafP) ~ SP + SITE,
        keep.order = TRUE),
  data = df_lmm |>
    drop_na())

df1 <- as_tibble(anova(mod1), rownames = "Predictor") |>
  select(Predictor, species_first=`Sum Sq`)

df1_sum <- df1 |>
  summarise(species_first = sum(species_first))

df1 <- df1 |>
  mutate(species_first_perc = species_first/df1_sum$species_first)
```

### Species last

```{r}
mod2 <- lm(
  terms(log(leafP) ~ SITE + SP,
        keep.order = TRUE),
  data = df_lmm |>
    drop_na())

df2 <- as_tibble(anova(mod2), rownames = "Predictor") |>
  select(Predictor, species_last=`Sum Sq`)

df2_sum <- df2 |>
  summarise(species_last = sum(species_last))

df2 <- df2 |>
  mutate(species_last_perc = species_last/df2_sum$species_last)
```

```{r}
df_site_species <- df1 |> 
  left_join(
    df2,
    by = "Predictor"
  )

df_site_species |> knitr::kable()
```

### Plot

```{r}
df_shared <- tibble(
  species = df_site_species |> 
    select(Predictor, species_last) |> 
    filter(Predictor == "SP") |> 
    pull(species_last),
  site = df_site_species |> 
    select(Predictor, species_first) |> 
    filter(Predictor == "SITE") |> 
    pull(species_first),
  species_plus = df_site_species |> 
    select(Predictor, species_first) |> 
    filter(Predictor == "SP") |> 
    pull(species_first),
  site_plus = df_site_species |> 
    select(Predictor, species_last) |> 
    filter(Predictor == "SITE") |> 
    pull(species_last)
) |> 
  mutate(
    shared_a = species_plus - species,
    shared_b = site_plus - site
  )

total <- df_shared$species_plus + df_shared$site

df_shared_perc <- df_shared |> 
  mutate(across(1:6, ~ 100*(. / total)))

df_venn <- tibble(x = c(0,9),
                  y = c(0,0),
                  r = c(sqrt(df_shared$site_plus), sqrt(df_shared$species_plus)),
                  id = c("site", "species")
                  )

gg_p <- ggplot(df_venn, 
             aes(x0 = x, y0 = y, r = r, fill = id)) +
  ggforce::geom_circle(alpha = .5, show.legend = FALSE ) +
  coord_fixed() +
  scale_fill_manual(values=c("#29a274ff", "#777055ff")) +
  theme_void() +
  annotate("text", x = -15 , y = 0, label = paste0("Site \n ", 
                          format(df_shared_perc$site, digits = 2), 
                          "%")) +
  annotate("text", x = 3 , y = 0, 
           label = paste0("Shared \n ", 
                          format(df_shared_perc$shared_a, digits = 2), 
                          "%")) +
  annotate("text", x = 24 , y = 0, 
           label = paste0("Species \n ", 
                          format(df_shared_perc$species, digits = 2), 
                          "%")) +
  labs(title = "Leaf P")

gg_p
```

## Venn diagram for leaf N:P

Here, site summarises all possible environmental factors.

### Species first

```{r}
mod1 <- lm(
  terms(log(LeafNP) ~ SP + SITE,
        keep.order = TRUE),
  data = df_lmm |>
    drop_na())

df1 <- as_tibble(anova(mod1), rownames = "Predictor") |>
  select(Predictor, species_first=`Sum Sq`)

df1_sum <- df1 |>
  summarise(species_first = sum(species_first))

df1 <- df1 |>
  mutate(species_first_perc = species_first/df1_sum$species_first)
```

### Species last

```{r}
mod2 <- lm(
  terms(log(LeafNP) ~ SITE + SP,
        keep.order = TRUE),
  data = df_lmm |>
    drop_na())

df2 <- as_tibble(anova(mod2), rownames = "Predictor") |>
  select(Predictor, species_last=`Sum Sq`)

df2_sum <- df2 |>
  summarise(species_last = sum(species_last))

df2 <- df2 |>
  mutate(species_last_perc = species_last/df2_sum$species_last)
```

```{r}
df_site_species <- df1 |> 
  left_join(
    df2,
    by = "Predictor"
  )

df_site_species |> knitr::kable()
```

### Plot

```{r}
df_shared <- tibble(
  species = df_site_species |> 
    select(Predictor, species_last) |> 
    filter(Predictor == "SP") |> 
    pull(species_last),
  site = df_site_species |> 
    select(Predictor, species_first) |> 
    filter(Predictor == "SITE") |> 
    pull(species_first),
  species_plus = df_site_species |> 
    select(Predictor, species_first) |> 
    filter(Predictor == "SP") |> 
    pull(species_first),
  site_plus = df_site_species |> 
    select(Predictor, species_last) |> 
    filter(Predictor == "SITE") |> 
    pull(species_last)
) |> 
  mutate(
    shared_a = species_plus - species,
    shared_b = site_plus - site
  )

total <- df_shared$species_plus + df_shared$site

df_shared_perc <- df_shared |> 
  mutate(across(1:6, ~ 100*(. / total)))

df_venn <- tibble(x = c(0,9),
                  y = c(0,0),
                  r = c(sqrt(df_shared$site_plus), sqrt(df_shared$species_plus)),
                  id = c("site", "species")
                  )

gg_np <- ggplot(df_venn, 
             aes(x0 = x, y0 = y, r = r, fill = id)) +
  ggforce::geom_circle(alpha = .5, show.legend = FALSE ) +
  coord_fixed() +
  scale_fill_manual(values=c("#29a274ff", "#777055ff")) +
  theme_void() +
  annotate("text", x = -10.5 , y = 0, label = paste0("Site \n ", 
                          format(df_shared_perc$site, digits = 2), 
                          "%")) +
  annotate("text", x = 3 , y = 0, 
           label = paste0("Shared \n ", 
                          format(df_shared_perc$shared_a, digits = 2), 
                          "%")) +
  annotate("text", x = 21 , y = 0, 
           label = paste0("Species \n ", 
                          format(df_shared_perc$species, digits = 2), 
                          "%")) +
  labs(title = "Leaf N:P")

gg_np
```

### Combined plot

```{r}
cowplot::plot_grid(gg_n, gg_p, gg_np, ncol = 3)
ggsave(paste0(here::here(), "/fig/venn.pdf"), width = 10, height = 3)
```
