---
title: "Trait gradient hypotheses"
description-title: "Supplementary Material to article: Balancing the power of environment versus phylogenetic controls on leaf nitrogen and phosphorous concentrations of terrestrial plants"
author: "Di Tian and Benjamin Stocker"
date: "2024-02-14"
format:
  pdf:
    toc: false
    number-sections: false
    colorlinks: true
---

## Biogeochemical Niche Hypothesis

The data-generating process is modeled following the Biogeochemical Niche Hypothesis as follows. A trait's value $x$ of observation $i$, of species $k$, at site $s$ is assumed to be exclusively determined by the identity of that species. All observed values of species $k$ are equal to that species' mean plus an observational error $\varepsilon_i$.

$$
x_{i,k,s} = \overline{x_k} + \varepsilon_i
$$
An environmental filtering on species occurrence is modeled as a random draw (sample with replacement) of $N_s$ individuals for each site from the population $N_k$ species. The probability to draw a species $k$ at site $s$ depends on the site and is modeled using a functional form as shown in @fig-prob. The "narrowness" of the the species distribution across different sites is controlled by a parameter $k$ (`param_species_selection`). 
```{r warning=FALSE, message=FALSE, echo=FALSE}
#| code-fold: true
#| label: fig-prob
#| fig-cap: "Probability for the occurrence of a given species at a given site."
#| out-width: 80%

library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(here)

get_prob <- function(
    isite, 
    nsites, 
    nspecies, 
    param_species_selection
    ){
  # preferred species at this site
  use_species <- min(max(floor(isite/nsites * nspecies), 1), nspecies)
  
  # get probabilities
  prob <- (abs((1:nspecies) - use_species) + 1)^param_species_selection

  return(prob)
}

nspecies <- 50

gg1 <- tibble(
  species = 1:nspecies,
  prob = get_prob(isite = 30, 
                  nsites = 100, 
                  nspecies = 50, 
                  param_species_selection = -2) 
) |> 
  ggplot(aes(species, prob)) +
  geom_point() +
  theme_classic() +
  labs(title = "Site number 30",
       x = "Species number",
       y = "Probability")

gg2 <- tibble(
  species = 1:nspecies,
  prob = get_prob(isite = 60, 
                  nsites = 100, 
                  nspecies = 50, 
                  param_species_selection = -2) 
) |> 
  ggplot(aes(species, prob)) +
  geom_point() +
  theme_classic() +
  labs(title = "Site number 60",
       x = "Species number",
       y = "Probability")

plot_grid(gg1, gg2)
```

The resulting trait gradient analysis (TGA) is shown in @fig-bgn. It demonstrates that the TGA identifies a data generating process following the Biogeochemical Niche Hypothesis by yielding  within-species regressions with a slope of zero.

```{r warning=FALSE, message=FALSE, echo=FALSE}
#| code-fold: true
#| label: fig-bgn
#| fig-cap: "Trait gradient analysis for data generated assuming the Biogeochemical Niche Hypothesis. Solid lines represent the species-specific regressions of trait values observed at different sites vs. the respective site mean across all species growing at that site."
#| out-width: 80%

nsites <- 100
nspecies <- 50
nobs_per_site <- 5
error <- 0
param_species_selection <- -2

set.seed(2)
df <- tibble()
idx <- 0
for (isite in 1:nsites){

  prob <- get_prob(isite, nsites, nspecies, param_species_selection)
    
  for (iobs in 1:nobs_per_site){
    idx <- idx + 1
    df <- bind_rows(
      df,
      tibble(
        idx = idx,
        site = isite,
        # At each site, species are randomly drawn from all available species 
        # with an uneven probability per species, proportional to 
        # [abs(species_index - preferred_species) + 1]^-2
        species = sample(1:nspecies, 
                         1, 
                         replace = TRUE, 
                         prob = prob
                         )
      ) |> 
        mutate(value = rnorm(1, mean = species, sd = error) )
    )
  }
}

df <- df |> 
  group_by(site) |> 
  summarise(sitemean = mean(value)) |> 
  right_join(
    df,
    by = "site"
  )

gg1 <- ggplot(data = df,
       aes(x = sitemean, y = value, group = species)) + 
  geom_smooth(method = "lm", 
              se = FALSE, 
              color = "black", 
              linewidth = 0.5) + 
  geom_point(color = "black", alpha = 0.4) +
  geom_abline(intercept = 0, 
              slope = 1, 
              linetype = "dotted", color = "red") +
  theme_classic() +
  labs(
    x = expression(paste("Site mean ", italic(x))),
    y = expression(paste("Observation ", italic(x))),
    title = "Biogeochemical Niche Hypothesis"
  )

gg1
```

```{r, echo=FALSE}
#| code-fold: true
#| label: fig-bgn-slopes
#| fig-cap: "Distribution of slopes in the trait gradient analysis for data generated assuming the Biogeochemical Niche Hypothesis."
#| out-width: 80%

df_tga <- df |> 
  group_by(species) |> 
  nest() |> 
  mutate(linmod = map(data, 
                      ~lm(value  ~ sitemean,  
                          data = ., 
                          x = TRUE))) |> 
  mutate(slope = map_dbl(linmod,  
                         ~coef(.)[2]))

gg2 <- df_tga |> 
  ggplot(aes(slope, after_stat(density))) + 
  geom_density( fill = "grey") +
  theme_classic() + 
  geom_vline(xintercept = 1.0, linetype = "dotted") + 
  geom_vline(xintercept = 0.0, linetype = "dotted") + 
  xlim(-0.5, 1.5) +
  labs(x = "Slope")

gg2
```


## Perfect stoichiometric plasticity

The data-generating process is modeled following perfect stoichiometric plasticit as follows. Species occur randomly across sites and the site's environmental condition imprints a common effect on all species growing at the site - a site-specific mean value $\overline{x}_x$. Species are characterized by a random offset ($\delta_k \sim N(0, \sigma)$) of the leaf trait relative to the mean across all species growing at site $s$. $\varepsilon_i$ is an observational error.

$$
x_{i,k,s} = \overline{x}_s + \delta_k + \varepsilon_i
$$
An environmental filtering on species occurrence is modeled as above, but with a slightly wider niche (smaller `param_species_selection`).

```{r warning=FALSE, message=FALSE, echo=FALSE}
#| code-fold: true
#| label: fig-prob_psp
#| fig-cap: "Probability for the occurrence of a given species at a given site. A wider niche is assumed here."
#| out-width: 80%

library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(here)

get_prob <- function(
    isite, 
    nsites, 
    nspecies, 
    param_species_selection
    ){
  # preferred species at this site
  use_species <- min(max(floor(isite/nsites * nspecies), 1), nspecies)
  
  # get probabilities
  prob <- (abs((1:nspecies) - use_species) + 1)^param_species_selection

  return(prob)
}

nspecies <- 50

gg1a <- tibble(
  species = 1:nspecies,
  prob = get_prob(isite = 30, 
                  nsites = 100, 
                  nspecies = 50, 
                  param_species_selection = -1) 
) |> 
  ggplot(aes(species, prob)) +
  geom_point() +
  theme_classic() +
  labs(title = "Site number 30",
       x = "Species number",
       y = "Probability")

gg2a <- tibble(
  species = 1:nspecies,
  prob = get_prob(isite = 60, 
                  nsites = 100, 
                  nspecies = 50, 
                  param_species_selection = -1) 
) |> 
  ggplot(aes(species, prob)) +
  geom_point() +
  theme_classic() +
  labs(title = "Site number 60",
       x = "Species number",
       y = "Probability")

plot_grid(gg1a, gg2a)
```

The resulting trait gradient analysis (TGA) is shown in @fig-psp. It demonstrates that the TGA identifies a data generating process following perfect stoichiometric plasticity by yielding  within-species regressions with a distribution of slopes that has a peak at one.

```{r warning=FALSE, message=FALSE, echo=FALSE}
#| code-fold: true
#| label: fig-psp
#| fig-cap: "Trait gradient analysis for data generated assuming perfect stoichiometric plasticity. Solid lines represent the species-specific regressions of trait values observed at different sites vs. the respective site mean across all species growing at that site."
#| out-width: 80%

nsites <- 100
nspecies <- 50
nobs_per_site <- 5
error <- 0
sd_species <- 15
param_species_selection <- -1

# generate the random offset per species
delta_species <- rnorm(nspecies, mean = 0, sd = sd_species)

set.seed(2)
df2 <- tibble()
idx <- 0
for (isite in 1:nsites){

  prob <- get_prob(isite, nsites, nspecies, param_species_selection)

  for (iobs in 1:nobs_per_site){
    idx <- idx + 1
    df2 <- bind_rows(
      df2,
      tibble(
        idx = idx,
        site = isite,
        # At each site, species are randomly drawn from all available species 
        # with an uneven probability per species, proportional to 
        # [abs(species_index - preferred_species) + 1]^-2
        species = sample(1:nspecies, 
                         1, 
                         replace = TRUE, 
                         prob = prob
                         )) |> 
        # site mean value taken equal to isite
        mutate(value = rnorm(1, mean = isite + delta_species[species], sd = error) )
    )
  }
}

df2 <- df2 |> 
  group_by(site) |> 
  summarise(sitemean = mean(value)) |> 
  right_join(
    df2,
    by = "site"
  )

gg3 <- ggplot(data = df2,
       aes(x = sitemean, y = value, group = species)) + 
  geom_smooth(method = "lm", 
              se = FALSE, 
              color = "black", 
              linewidth = 0.5) + 
  geom_point(color = "black", alpha = 0.4) +
  geom_abline(intercept = 0, 
              slope = 1, 
              linetype = "dotted", color = "red") +
  theme_classic() +
  labs(
    x = expression(paste("Site mean ", italic(x))),
    y = expression(paste("Observation ", italic(x))),
    title = "Perfect stoichiometric plasticity"
  )

gg3
```

```{r, echo=FALSE}
#| code-fold: true
#| label: fig-psp-slopes
#| fig-cap: "Distribution of slopes in the trait gradient analysis for data generated assuming the perfect stoichiometric plasticity."
#| out-width: 80%

df_tga2 <- df2 |> 
  group_by(species) |> 
  nest() |> 
  mutate(linmod = map(data, 
                      ~lm(value  ~ sitemean,  
                          data = ., 
                          x = TRUE))) |> 
  mutate(slope = map_dbl(linmod,  
                         ~coef(.)[2]))

gg4 <- df_tga2 |> 
  ggplot(aes(slope, after_stat(density))) + 
  geom_density( fill = "grey") +
  theme_classic() + 
  geom_vline(xintercept = 1.0, linetype = "dotted") + 
  geom_vline(xintercept = 0.0, linetype = "dotted") + 
  xlim(-0.5, 1.5) +
  labs(x = "Slope")

gg4
```

## Publication figure

```{r, echo=FALSE}
# plot_grid(gg1, gg3, ncol = 2)
# ggsave(here("fig/tga_demo.pdf"), width = 8, height = 4)
# ggsave(here("fig/tga_demo.png"), width = 8, height = 4)

plot_grid(gg1, gg3, gg2, gg4, nrow = 2, rel_heights = c(1, 0.7))
ggsave(here("fig/tga_demo_slopes.pdf"), width = 8, height = 6)
ggsave(here("fig/tga_demo_slopes.png"), width = 8, height = 6)
```