---
title: "Trait gradient analysis"
author: "Beni"
date: "7/6/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(RColorBrewer)
```

## Prepare data
```{r}
df <- read_csv("~/data/LeafNP_tiandi/Global_total_leaf_N_P_Di/leafnp_data_covariates_20210702.csv") %>% 
  dplyr::select(leafN, leafP, LeafNP, Family, Genus, Species, sitename) |> 
  
  # correct
  mutate(Family = ifelse(Species == "Acer saccharum", "Sapindaceae", Family))
```

Add logarithms
```{r}
# Had it like this first:
# df <- df %>% 
#   mutate(log_leafN = log10(leafN), 
#          log_leafP = log10(leafP),
#          log_leafN_sitemean = log10(leafN_sitemean),
#          log_leafP_sitemean = log10(leafP_sitemean)
#          )

df <- df %>%
  mutate(log_leafN  = log10(leafN),
         log_leafP  = log10(leafP),
         log_leafNP = log10(LeafNP))

df <- df |> 
  group_by(sitename) |> 
  summarise(log_leafN_sitemean  = mean(log_leafN),
            log_leafP_sitemean  = mean(log_leafP),
            log_leafNP_sitemean = mean(log_leafNP),
            leafN_sitemean  = mean(leafN), 
            leafP_sitemean  = mean(leafP),
            leafNP_sitemean = mean(LeafNP)) %>% 
  
  # Add site-mean to full data.
  right_join(df, by = "sitename")
```

Determine most common species (for plotting) with more than 50 observations
```{r}
# determine most common species
common_species <- df |> 
  group_by(Species) |> 
  summarise(n = n()) |> 
  filter(n >= 50) |> 
  pull(Species)
```

## Filter data

Use data only for species that appear in at least N sites.
```{r}
use_species <- df %>% 
  dplyr::select(sitename, Species) %>% 
  distinct() %>% 
  group_by(Species) %>% 
  summarise(n = n()) %>% 
  dplyr::filter(n >= 5) %>% 
  pull(Species)
```

Use data only for sites where at least M species are present.
```{r}
use_sites <- df %>% 
  dplyr::select(sitename, Species) %>% 
  distinct() %>% 
  group_by(sitename) %>% 
  summarise(n = n()) %>% 
  dplyr::filter(n >= 5) %>% 
  pull(sitename)
```

## TGA Plot

Plot.
```{r eval=FALSE}
df %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  ggplot(aes(x = leafN_sitemean, y = leafN, color = Species)) +
  geom_point()
```

### Only lines

Plot just the lines
```{r}
# N
df %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  group_by(Species) %>% 
  ggplot(aes(x = leafN_sitemean, y = leafN, group = Species)) +
  geom_smooth(method = "lm", se = FALSE, size = 0.1, alpha = 0.2) +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  coord_equal() +
  theme_classic()
  # geom_point(alpha = 0.3)

# P
df %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  group_by(Species) %>% 
  ggplot(aes(x = leafP_sitemean, y = leafP, group = Species)) +
  geom_smooth(method = "lm", se = FALSE, size = 0.1, alpha = 0.2) +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  coord_equal() +
  theme_classic()
  # geom_point(alpha = 0.3)

# N:P
df %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  group_by(Species) %>% 
  ggplot(aes(x = leafNP_sitemean, y = LeafNP, group = Species)) +
  geom_smooth(method = "lm", se = FALSE, size = 0.1, alpha = 0.2) +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  coord_equal() +
  theme_classic()
  # geom_point(alpha = 0.3)
```

### Analysis of slopes

Fit linear regressions by species.
```{r}
df_tga <- df %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  group_by(Species) %>% 
  nest() %>% 
  mutate(linmod_n  = purrr::map(data, ~lm(leafN  ~ leafN_sitemean,  data = ., x = TRUE)),
         linmod_p  = purrr::map(data, ~lm(leafP  ~ leafP_sitemean,  data = ., x = TRUE)),
         linmod_np = purrr::map(data, ~lm(LeafNP ~ leafNP_sitemean, data = ., x = TRUE))) %>% 
  mutate(slope_n  = purrr::map_dbl(linmod_n,  ~coef(.)[2]),
         slope_p  = purrr::map_dbl(linmod_p,  ~coef(.)[2]),
         slope_np = purrr::map_dbl(linmod_np, ~coef(.)[2]),
         range_n  = purrr::map(linmod_n,  "x"),
         range_p  = purrr::map(linmod_p,  "x"),
         range_np = purrr::map(linmod_np, "x")) %>% 
  mutate(range_n  = purrr::map_dbl(range_n,  ~diff(range(.))),
         range_p  = purrr::map_dbl(range_p,  ~diff(range(.))),
         range_np = purrr::map_dbl(range_np, ~diff(range(.)))) |> 
  left_join(df %>% 
              dplyr::select(Family, Genus, Species) %>% 
              distinct(), 
            by = "Species")

df_tga %>% 
  ggplot() +
  geom_histogram(aes(slope_n, ..density..), fill = "royalblue", binwidth = 0.1, alpha = 0.5) +
  geom_density(aes(slope_n, ..density..), color = "royalblue") +
  geom_histogram(aes(slope_p, ..density..), fill = "tomato", binwidth = 0.1, alpha = 0.5) +
  geom_density(aes(slope_p, ..density..), color = "tomato") +
  xlim(-2,3) + 
  theme_classic()

df_tga %>% 
  group_by(Family) %>% 
  summarise(slope_n = mean(slope_n)) %>% 
  mutate(Family = fct_reorder(Family, slope_n)) %>% 
  drop_na() %>% 
  ggplot(aes(Family, slope_n)) +
  geom_bar(stat = "identity") +
  coord_flip()

```

## Logarithmic TGA Plot

Plot.
```{r eval=FALSE}
df %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  ggplot(aes(x = log_leafN_sitemean, y = log_leafN, color = Species)) +
  geom_point()
```

### Only lines

Plot just the lines
```{r}
highcontrast <- khroma::colour("high contrast")
cols <- highcontrast(3)

# N
gg1 <- df %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  drop_na(log_leafN_sitemean, log_leafN) |> 
  group_by(Species) %>% 
  ggplot(aes(x = log_leafN_sitemean, y = log_leafN, group = Species)) +
  geom_smooth(method = "lm", se = FALSE, size = 0.1, alpha = 0.2, colour = cols[1]) +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  theme_classic() +
  xlim(0.7, 1.8) + ylim(0.7, 1.8) +
  labs(x = expression(paste("site mean of log"[10],"(Leaf N)")), y = expression(paste("log"[10], "(Leaf N)")), title = "Leaf N")
  # geom_point(alpha = 0.3)

gg1
ggsave(paste0(here::here(), "/fig/tga_leafn_log.pdf"), width = 3.5, height = 4)

# P
gg2 <- df %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  group_by(Species) %>% 
  ggplot(aes(x = log_leafP_sitemean, y = log_leafP, group = Species)) +
  geom_smooth(method = "lm", se = FALSE, size = 0.1, alpha = 0.2, colour = cols[2]) +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  xlim(-0.6, 0.8) + ylim(-0.6, 0.8) +
  theme_classic() + 
  labs(x = expression(paste("site mean of log"[10],"(Leaf P)")), y = expression(paste("log"[10], "(Leaf P)")), title = "Leaf P")
  # geom_point(alpha = 0.3)

gg2
ggsave(paste0(here::here(), "/fig/tga_leafp_log.pdf"), width = 3.5, height = 4)

# N:P
gg3 <- df %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  group_by(Species) %>% 
  ggplot(aes(x = log_leafNP_sitemean, y = log_leafNP, group = Species)) +
  geom_smooth(method = "lm", se = FALSE, size = 0.1, alpha = 0.2, colour = cols[3]) +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  xlim(0.5, 2) + ylim(0.5, 2) +
  theme_classic() + 
  labs(x = expression(paste("site mean of log"[10],"(Leaf N:P)")), y = expression(paste("log"[10], "(Leaf N:P)")), title = "Leaf N:P")
  # geom_point(alpha = 0.3)

gg3
ggsave(paste0(here::here(), "/fig/tga_leafnp_log.pdf"), width = 3.5, height = 4)

toprow <- cowplot::plot_grid(gg1, gg2, gg3, labels = c("a", "b", "c"), ncol = 3)
toprow
ggsave(paste0(here::here(), "/fig/tga_log.pdf"), width = 12, height = 4)
```

### Analysis of slopes

Fit linear regressions by species.
```{r}
get_x <- function(linmod){
  as.vector(linmod$x[,2])
}
         
df_agg <- df |> 
  ungroup() |> 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) |> 
  nest(data = everything()) |> 
  mutate(range_n     = purrr::map_dbl(data, ~diff(quantile(pull(., leafN), probs = c(0.01, 0.99)))),
         range_p     = purrr::map_dbl(data, ~diff(quantile(pull(., leafP), probs = c(0.01, 0.99)))),
         range_np    = purrr::map_dbl(data, ~diff(quantile(pull(., LeafNP), probs = c(0.01, 0.99))))
         ) |> 
  select( -data )

# df_agg_species <- df |> 
#   
#   # subset
#   dplyr::filter(Species %in% use_species & sitename %in% use_sites) |> 
#   
#   # fit linear regressions by species
#   group_by(Species) |> 
#   nest() |> 
#   mutate(leafN_mean  = purrr::map_dbl(data, ~mean(pull(., leafN))),
#          leafP_mean  = purrr::map_dbl(data, ~mean(pull(., leafP))),
#          leafNP_mean = purrr::map_dbl(data, ~mean(pull(., LeafNP))),
#          range_n     = purrr::map_dbl(data, ~diff(quantile(pull(., leafN), probs = c(0.01, 0.99)))),
#          range_p     = purrr::map_dbl(data, ~diff(quantile(pull(., leafP), probs = c(0.01, 0.99)))),
#          range_np    = purrr::map_dbl(data, ~diff(quantile(pull(., LeafNP), probs = c(0.01, 0.99))))
#          ) |> 
#   mutate(range_n_norm = range_n / df_agg$range_n,
#          range_p_norm = range_p / df_agg$range_p,
#          range_np_norm = range_np / df_agg$range_np,
#          )

df_tga <- df |> 
  
  # subset
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  
  # fit linear regressions by species
  group_by(Species) %>% 
  nest() %>% 
  mutate(linmod_log_n  = purrr::map(data, ~lm(log_leafN  ~ log_leafN_sitemean,  data = ., x = TRUE)),
         linmod_log_p  = purrr::map(data, ~lm(log_leafP  ~ log_leafP_sitemean,  data = ., x = TRUE)),
         linmod_log_np = purrr::map(data, ~lm(log_leafNP ~ log_leafNP_sitemean, data = ., x = TRUE)),
         
         linmod_n  = purrr::map(data, ~lm(leafN  ~ leafN_sitemean,  data = ., x = TRUE)),
         linmod_p  = purrr::map(data, ~lm(leafP  ~ leafP_sitemean,  data = ., x = TRUE)),
         linmod_np = purrr::map(data, ~lm(LeafNP ~ leafNP_sitemean, data = ., x = TRUE)),
         ) %>% 
  
  # slope in log-TGA
  mutate(slope_n  = purrr::map_dbl(linmod_log_n,  ~coef(.)[2]),
         slope_p  = purrr::map_dbl(linmod_log_p,  ~coef(.)[2]),
         slope_np = purrr::map_dbl(linmod_log_np, ~coef(.)[2]),
         
         # site-level mean values from (raw) TGA
         range_n  = purrr::map(linmod_n,  ~get_x(.)),
         range_p  = purrr::map(linmod_p,  ~get_x(.)),
         range_np = purrr::map(linmod_np, ~get_x(.))) |> 
  
  # range in site-level mean from raw TGA
  mutate(range_n  = purrr::map_dbl(range_n,  ~diff(quantile(., probs = c(0.01, 0.99)))),
         range_p  = purrr::map_dbl(range_p,  ~diff(quantile(., probs = c(0.01, 0.99)))),
         range_np = purrr::map_dbl(range_np, ~diff(quantile(., probs = c(0.01, 0.99)))),
         ) |> 
  
  mutate(range_n_norm = range_n / df_agg$range_n,
         range_p_norm = range_p / df_agg$range_p,
         range_np_norm = range_np / df_agg$range_np,
         ) |> 
  
  # add phylo info
  left_join(df|> 
              dplyr::select(Family, Genus, Species)|> 
              distinct(), 
            by = "Species") |> 
  
  # remove outlier in slope P
  filter(slope_p > -5 & slope_p < 5 & range_p_norm > 0)
```

Maximum density (mode) of the distribution of slopes.
```{r}
vec <- df_tga$slope_n

get_mode <- function(vec){
  # based on https://stackoverflow.com/questions/58785930/r-find-maximum-of-density-plot
 
  # exclude wiggles on either end
  vec <- vec[vec > -0.5 & vec < 2]
 
  d <- density(vec)
  i <- which(diff(sign(diff(d$y))) < 0) + 1
  out <- d$x[i]
  
  # out <- data.frame(x = d$x[i], y = d$y[i])  
  return(out)
}

df_modes <- df_tga |> 
  ungroup() |> 
  summarise(mode_n  = get_mode(slope_n),
            mode_p  = get_mode(slope_p),
            mode_np = get_mode(slope_np)
            )

df_modes
```


```{r}
# distribution of slopes
gg_slopes <- df_tga|> 
  pivot_longer(c(slope_n, slope_p, slope_np), values_to = "slope", names_to = "trait") |> 
  mutate(trait = factor(trait, level = c("slope_n", "slope_p", "slope_np"))) |> 
  ggplot(aes(slope, ..density.., color = trait, fill = trait)) +
  geom_density(alpha = 0.3) +
  khroma::scale_color_highcontrast(labels = c("N", "P", "N:P")) +
  khroma::scale_fill_highcontrast(labels = c("N", "P", "N:P")) +
  xlim(-2,3) +
  theme_classic() +
  labs(x = "Slope", y = "Density") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  theme(legend.title = element_blank())

gg_slopes
ggsave(paste0(here::here(), "/fig/tga_slopes_density.pdf"), width = 7, height = 5)

# distribution of ranges
df_tga_median <- df_tga|> 
  pivot_longer(c(range_n_norm, range_p_norm, range_np_norm), 
               values_to = "range", 
               names_to = "trait") |>
  group_by(trait) |> 
  summarise(across(range, median))

df_tmp <- df_tga %>% 
  pivot_longer(c(range_n_norm, range_p_norm, range_np_norm), 
               values_to = "range", 
               names_to = "trait") |> 
  mutate(trait = factor(trait, level = c("range_n_norm", "range_p_norm", "range_np_norm")))

gg_ranges <- ggplot() +
  geom_density(data = df_tmp, 
               aes(range, ..density.., color = trait, fill = trait), 
               alpha = 0.3) +
  geom_vline(data = df_tga_median, 
             aes(xintercept = range, color = trait)) +
  khroma::scale_color_highcontrast(labels = c("N", "P", "N:P")) +
  khroma::scale_fill_highcontrast(labels = c("N", "P", "N:P")) +
  # xlim(-2,3) +
  theme_classic() +
  labs(x = "Normalised range", y = "Density") +
  theme(legend.title = element_blank())

gg_ranges
ggsave(paste0(here::here(), "/fig/tga_ranges_density.pdf"), width = 7, height = 5)

# mean slopes by family: N
df_tga |> 
  group_by(Family) %>% 
  summarise(slope_n = mean(slope_n)) %>% 
  mutate(Family = fct_reorder(Family, slope_n)) %>% 
  drop_na() %>% 
  ggplot(aes(Family, slope_n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_classic() +
  labs(y = "Slope", title = "Leaf N") +
  geom_hline(yintercept = 1.0, linetype = "dotted") +
  geom_hline(yintercept = 0.0) +
  khroma::scale_fill_discreterainbow()

ggsave(paste0(here::here(), "/fig/tga_slopes_bars_leafn.pdf"), width = 7, height = 10)

# mean slopes by family: P
df_tga|> 
  group_by(Family) %>% 
  summarise(slope_p = mean(slope_p)) %>% 
  mutate(Family = fct_reorder(Family, slope_p)) %>% 
  drop_na() %>% 
  ggplot(aes(Family, slope_p)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_classic() +
  labs(y = "Slope", title = "Leaf P") +
  geom_hline(yintercept = 1.0, linetype = "dotted") +
  geom_hline(yintercept = 0.0) +
  khroma::scale_fill_discreterainbow()

ggsave(paste0(here::here(), "/fig/tga_slopes_bars_leafp.pdf"), width = 7, height = 10)

# mean slopes of most common species: N
gg_slopes_species_n <- df_tga|> 
  filter(Species %in% common_species) |> 
  group_by(Species) %>%
  summarise(slope_n = mean(slope_n)) %>%
  ungroup() |> 
  drop_na() %>% 
  left_join(df_tga |> 
              select(Species, Family, Genus) |> 
              distinct(),
            by = "Species") |> 
  mutate(Species = fct_reorder(Species, slope_n)) %>% 
  ggplot(aes(Species, slope_n, fill = Family)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  theme_classic() +
  labs(y = "Slope", title = "Leaf N") +
  geom_hline(yintercept = 1.0, linetype = "dotted") +
  geom_hline(yintercept = 0.0) +
  # scale_fill_manual(values = as.vector(pals::stepped(18)))
  khroma::scale_fill_discreterainbow() +
  theme(axis.text.y = element_text(face = "italic"))

gg_slopes_species_n
ggsave(paste0(here::here(), "/fig/tga_slopes_bars_leafn_species.pdf"), width = 7, height = 7)

# mean slopes of most common species: P
gg_slopes_species_p <- df_tga |> 
  filter(Species %in% common_species) |> 
  group_by(Species) |>
  summarise(slope_p = mean(slope_p)) |>
  ungroup() |> 
  drop_na() |> 
  left_join(df_tga |> 
              select(Species, Family, Genus) |> 
              distinct(),
            by = "Species") |> 
  mutate(Species = fct_reorder(Species, slope_p)) |> 
  ggplot(aes(Species, slope_p, fill = Family)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  theme_classic() +
  labs(y = "Slope", title = "Leaf P") +
  geom_hline(yintercept = 1.0, linetype = "dotted") +
  geom_hline(yintercept = 0.0) +
  khroma::scale_fill_discreterainbow() +
  theme(axis.text.y = element_text(face = "italic"))

gg_slopes_species_p
ggsave(paste0(here::here(), "/fig/tga_slopes_bars_leafp_species.pdf"), width = 7, height = 7)

# mean slopes of most common species: N:P
gg_slopes_species_np <- df_tga |> 
  filter(Species %in% common_species) |> 
  group_by(Species) |>
  summarise(slope_np = mean(slope_np)) |>
  ungroup() |> 
  drop_na() |> 
  left_join(df_tga |> 
              select(Species, Family, Genus) |> 
              distinct(),
            by = "Species") |> 
  mutate(Species = fct_reorder(Species, slope_np)) |> 
  ggplot(aes(Species, slope_np, fill = Family)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_classic() +
  labs(y = "Slope", title = "Leaf N:P") +
  geom_hline(yintercept = 1.0, linetype = "dotted") +
  geom_hline(yintercept = 0.0) +
  khroma::scale_fill_discreterainbow() +
  theme(axis.text.y = element_text(face = "italic"))

gg_slopes_species_np
ggsave(paste0(here::here(), "/fig/tga_slopes_bars_leafnp_species.pdf"), width = 7, height = 7)
```

Slopes by family
```{r}
df_families <- df_tga |> 
  group_by(Family) |> 
  summarise(n = n()) |> 
  arrange(desc(n)) |> 
  filter(n >= 5)
  
use_families <- df_families |> 
  pull(Family)

# Leaf N
gg1 <- df_tga |> 
  ungroup() |> 
  filter(Family %in% use_families) |> 
  left_join(df_families, by = "Family") |> 
  mutate(Family = paste0(Family, " (", as.character(n), ")")) |> 
  mutate(Family = fct_reorder(Family, slope_n)) |> 
  ggplot(aes(x = slope_n, y = Family)) +
  ggridges::geom_density_ridges(scale = 2) + 
  xlim(-2, 3) +
  coord_cartesian(clip = "off") + # to avoid clipping of the very top of the top ridgeline
  ggridges::theme_ridges() +
  geom_vline(xintercept = c(0,1), linetype = "dotted") +
  labs(title = "Leaf N", x = "Slope")

# Leaf P
gg2 <- df_tga |> 
  ungroup() |> 
  filter(Family %in% use_families) |> 
  left_join(df_families, by = "Family") |> 
  mutate(Family = paste0(Family, " (", as.character(n), ")")) |> 
  mutate(Family = fct_reorder(Family, slope_p)) |> 
  ggplot(aes(x = slope_p, y = Family)) +
  ggridges::geom_density_ridges(scale = 2) + 
  xlim(-2, 3) +
  coord_cartesian(clip = "off") + # to avoid clipping of the very top of the top ridgeline
  ggridges::theme_ridges() +
  geom_vline(xintercept = c(0,1), linetype = "dotted") +
  labs(title = "Leaf P", x = "Slope")

# Leaf N:P
gg3 <- df_tga |> 
  ungroup() |> 
  filter(Family %in% use_families) |> 
  left_join(df_families, by = "Family") |> 
  mutate(Family = paste0(Family, " (", as.character(n), ")")) |> 
  mutate(Family = fct_reorder(Family, slope_n)) |> 
  ggplot(aes(x = slope_n, y = Family)) +
  ggridges::geom_density_ridges(scale = 2) + 
  xlim(-2, 3) +
  coord_cartesian(clip = "off") + # to avoid clipping of the very top of the top ridgeline
  ggridges::theme_ridges() +
  geom_vline(xintercept = c(0,1), linetype = "dotted") +
  labs(title = "Leaf N:P", x = "Slope")

plot_grid(gg1, gg2, gg3, nrow = 1)
ggsave(paste0(here::here(), "/fig/slopes_families.pdf"), width = 12, height = 5)
```

```{r}
median(df_tga$range_n_norm)
median(df_tga$range_p_norm)
median(df_tga$range_np_norm)
```

### Correlations of slopes and ranges

```{r}
# correlation of slopes for N and P
out <- df_tga |> 
  # remove outliers
  filter(slope_n > -5, slope_n < 5, slope_p > -5, slope_p < 5) |> 
  rbeni::analyse_modobs2("slope_p", "slope_n", rsquared = FALSE)

out$gg

# correlation of slopes and range for N
out <- df_tga |> 
  # remove outliers
  filter(slope_n > -5, slope_n < 5, slope_p > -5, slope_p < 5) |> 
  rbeni::analyse_modobs2("slope_n", "range_n", rsquared = FALSE)

out$gg

# correlation of slopes and range for P
out <- df_tga |> 
  # remove outliers
  filter(slope_n > -5, slope_n < 5, slope_p > -5, slope_p < 5) |> 
  rbeni::analyse_modobs2("slope_p", "range_p", rsquared = FALSE)

out$gg

# correlation of ranges for N and P
r_pearson <- format(cor(df_tga$range_n, df_tga$range_p), digits = 2)
df_tga |> 
  # remove outliers
  filter(slope_n > -5, slope_n < 5, slope_p > -5, slope_p < 5) |> 
  ggplot(aes(range_n, range_p)) +
  geom_point() +
  labs(x = expression(paste("Leaf N range (g N (g DM)"^-1, ")")),
       y = expression(paste("Leaf P range (g P (g DM)"^-1, ")")),
       subtitle = bquote( italic(r) == .(r_pearson) )
      ) +
  theme_classic()
ggsave(paste0(here::here(), "/fig/cor_ranges.pdf"), width = 5, height = 5)
ggsave(paste0(here::here(), "/fig/cor_ranges.png"), width = 5, height = 5)
```

## Species variation decomposition

Make two alternative (extreme) assumptions and test how much of the overall variations $x_i$ these simplifying scenarios explain:

- Scenario 1: All variations are across species. Species all have their fixed $\overline{x_{i}}$.
- Scenario 2: All variations are within species, no variations between species. All species have the same mean $\overline{x}$. 

```{r}
df_species <- df |>
  
  # subset
  dplyr::select(all_of(c("sitename", "Species", "leafN", "leafP", "LeafNP"))) |> 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) |> 
  
  # fit linear regressions by species
  group_by(Species) |> 
  summarise(across(all_of(c("leafN", "leafP", "LeafNP")), mean)) |> 
  rename(leafN_mean = leafN, leafP_mean = leafP, leafNP_mean = LeafNP)

df_species_mean <- df_species |> 
  ungroup() |> 
  summarise(across(all_of(c("leafN_mean", "leafP_mean", "leafNP_mean")), mean))
  
df2 <- df |> 

  # subset
  dplyr::select(all_of(c("sitename", "Species", "leafN", "leafP", "LeafNP"))) |> 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) |> 
  
  left_join(df_species, by = "Species") |> 
  ungroup() |> 
  mutate(leafN_diff = leafN - leafN_mean, 
         leafP_diff = leafP - leafP_mean, 
         leafNP_diff = LeafNP - leafNP_mean) |> 
  
  # H1: All variations are among species. Species all have their fixed $\overline{x_{i}}$.
  mutate(leafN_h1  = leafN_mean,
         leafP_h1  = leafP_mean,
         leafNP_h1 = leafNP_mean) |> 
  
  # H2: All variations are within species, no variations between species. 
  # All species have the same mean $\overline{x}$.
  mutate(leafN_h2  = df_species_mean$leafN_mean  + leafN_diff,
         leafP_h2  = df_species_mean$leafP_mean  + leafP_diff,
         leafNP_h2 = df_species_mean$leafNP_mean + leafNP_diff,
         )
  
# collect R-squared of the partitionings
df_part <- tibble() |> 
  
  # leaf N
  bind_rows(
    yardstick::rsq_trad(df2, "leafN", "leafN_h1") |> 
      mutate(var = "leafN", type = "across"), 
  ) |> 
  bind_rows(
    yardstick::rsq_trad(df2, "leafN", "leafN_h2") |> 
      mutate(var = "leafN", type = "within"), 
  ) |> 
  
  # leaf P
  bind_rows(
    yardstick::rsq_trad(df2, "leafP", "leafP_h1") |> 
      mutate(var = "leafP", type = "across"), 
  ) |> 
  bind_rows(
    yardstick::rsq_trad(df2, "leafP", "leafP_h2") |> 
      mutate(var = "leafP", type = "within"), 
  ) |> 
  
  # leaf P
  bind_rows(
    yardstick::rsq_trad(df2, "LeafNP", "leafNP_h1") |> 
      mutate(var = "leafNP", type = "across"), 
  ) |> 
  bind_rows(
    yardstick::rsq_trad(df2, "LeafNP", "leafNP_h2") |> 
      mutate(var = "leafNP", type = "within"), 
  )
  # pivot_wider(names_from = "type", values_from = ".estimate")

gg_part <- df_part |>  
  mutate(var = ifelse(var == "leafN", "Leaf N",
                      ifelse(var == "leafP", "Leaf P",
                             "Leaf N:P"))) |> 
  mutate(var = factor(var, levels = rev(c("Leaf N", "Leaf P", "Leaf N:P")))) |> 
  
  ggplot(aes(x = var, y = .estimate, fill = type)) +
  geom_bar(stat="identity", position = "fill", width = 0.6) +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(values = c("#777055ff", "#29a274ff"),
                    labels = c("across", "within"),
                    name = NULL) +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  labs(title = "Species variation decomposition",
       y = "Fraction of variation explained",
       x = "")
```


## Log TGA with RF predictions

Add predictions from the RF model. Column `pred` contains predictions.
```{r}
filn <- paste0(here::here(), "/data/mod_rf_caret_leafn.rds")  # created with model_fitting.Rmd
mod_rf_caret_leafn <- readRDS(filn)

filn <- paste0(here::here(), "/data/dfs_leafnp_20210729.rds")
dfs <- readRDS(filn)

# get predicted values from cross-validation resamples
dfs_pred <- mod_rf_caret_leafn$pred %>% 
  as_tibble() %>% 
  dplyr::filter(mtry == mod_rf_caret_leafn$bestTune$mtry, 
                splitrule == mod_rf_caret_leafn$bestTune$splitrule, 
                min.node.size == mod_rf_caret_leafn$bestTune$min.node.size) %>%
  separate(Resample, into = c(NA, "Fold"), sep = "old") %>% 
  dplyr::rename(idx = rowIndex) |> 
  
  # combine to get site name
  left_join(
    dfs |> 
      mutate(idx = 1:nrow(dfs)),
    by = "idx"
    ) |> 
  
  # retain only site name and predicted
  select(sitename, pred)

# add predicted values to full data frame
df_rf <- df |> 
  left_join(
    dfs_pred,
    by = "sitename"
  )
  
```

Treat `pred` like site-mean values in "classic" TGA.
```{r}
# N
gg1 <- df_rf %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  drop_na(leafN_sitemean, pred) |> 
  group_by(Species) %>% 
  ggplot(aes(x = pred, y = leafN, group = Species)) +
  geom_smooth(method = "lm", se = FALSE, size = 0.1, alpha = 0.2, colour = cols[1]) +
  geom_abline(intercept=0, slope=1, linetype="dotted") +
  theme_classic() +
  # xlim(0.7, 1.8) + ylim(0.7, 1.8) +
  # coord_equal() +
  labs(x = expression(paste("Predicted leaf N")), y = expression(paste("Leaf N")), title = "Leaf N")
gg1
```

### Analysis of slopes

Fit linear regressions by species.
```{r}
df_tga2 <- df_rf %>% 
  dplyr::filter(Species %in% use_species & sitename %in% use_sites) %>% 
  group_by(Species) %>% 
  nest() %>% 
  mutate(linmod_n  = purrr::map(data, ~lm(leafN  ~ pred,  data = ., x = TRUE))
         # linmod_p  = purrr::map(data, ~lm(log_leafP  ~ log_leafP_sitemean,  data = ., x = TRUE)),
         # linmod_np = purrr::map(data, ~lm(log_leafNP ~ log_leafNP_sitemean, data = ., x = TRUE))
         ) %>% 
  mutate(slope_n  = purrr::map_dbl(linmod_n,  ~coef(.)[2]),
         # slope_p  = purrr::map_dbl(linmod_p,  ~coef(.)[2]),
         # slope_np = purrr::map_dbl(linmod_np, ~coef(.)[2]),
         
         range_n  = purrr::map(linmod_n,  "x")
         # range_p  = purrr::map(linmod_p,  "x"),
         # range_np = purrr::map(linmod_np, "x")
         ) %>% 
  mutate(range_n  = purrr::map_dbl(range_n,  ~diff(range(.)))
         # range_p  = purrr::map_dbl(range_p,  ~diff(range(.))),
         # range_np = purrr::map_dbl(range_np, ~diff(range(.)))
         ) |> 
  left_join(df_rf %>% 
              dplyr::select(Family, Genus, Species) %>% 
              distinct(), 
            by = "Species")
  
  # # remove outlier in slope P
  # filter(slope_p > -5 & slope_p < 5)
```


```{r}
df_tga2 %>% 
  # pivot_longer(c(slope_n, slope_p, slope_np), values_to = "slope", names_to = "trait") |> 
  # mutate(trait = factor(trait, level = c("slope_n", "slope_p", "slope_np"))) |> 
  ggplot(aes(slope_n, ..density..)) +   # , color = trait, fill = trait
  geom_density(alpha = 0.3) +
  # khroma::scale_color_highcontrast(labels = c("N", "P", "N:P")) +
  # khroma::scale_fill_highcontrast(labels = c("N", "P", "N:P")) +
  xlim(-2,3) +
  theme_classic() +
  labs(x = "Slope", y = "Density") +
  geom_vline(xintercept = 1, linetype = "dotted") +
  theme(legend.title = element_blank())
```


## Publication figure

```{r}
midrow <- cowplot::plot_grid(gg_slopes, gg_ranges, gg_part, labels = c("d", "e", "f"), ncol = 3)
midrow
ggsave(paste0(here::here(), "/fig/tga_log_mid.pdf"), width = 12, height = 4)

bottomrow <- cowplot::plot_grid(gg_slopes_species_n, gg_slopes_species_p, gg_slopes_species_np, labels = c("g", "h", "i"), ncol = 3, rel_widths = c(0.65, 0.65, 1))
bottomrow
ggsave(paste0(here::here(), "/fig/tga_log_bottom.pdf"), width = 12, height = 5)

plot_grid(toprow, midrow, bottomrow, nrow = 3, rel_heights = c(0.7, 0.5, 1.1))
ggsave(paste0(here::here(), "/fig/tga_log_all.pdf"), width = 12, height = 13)
```